---
title: "MATH219 Lab4"
author: "Leanne Dong"
subtitle: "Simulation, Discrete distribution"
date: "18/08/2018"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

# Simulation (by hand)

Q1: Attempt the questions leftover from Lecture 3 extra exercises on simulation.

# Simulation (in R)

## Example: Tossing a coin

As you already seen, the theoretical probability of getting a head or tail on a coin toss is both 0.5. Now Let us use R to simulate an experiment of flipping a coin a number of times and compare our results with the theoretical probability. First let fix the convention:

**0=Tail and 1=Heads**

We can use the following command to ask R to flip a coin 10 times:

```{r}
sample(0:1,10,rep=T)
```
This gives 4 tails and 6 head. In fact we can write a function to flip a coin $n$ times:

```{r}
FlipCoin=function(n) sample(0:1,n,rep=T)
e1=FlipCoin(50)
e1
```
Now we can use the sum command to compare the results from this experiement to the theoretical probabilities. For instance, in the above experiment of flipping a coin 50 times, we can count the heads and tails as

```{r}
sum(e1==0)
sum(e1==0)/50
sum(e1==1)
sum(e1==1)/50
```
This gives us 24 tails and 26 heads. The "probability" of relative frequency of a Tail in this experiment is 0.48 and a Head is 0.52. (Note, you may get different answers from mine. Think why!)
Let us now plot a relative histogram:
```{r}
hist(e1,breaks=c(-0.5,0.5,1.5),prob=T)
```

## Rolling dice

The probability of getting a number between 1 to 6 on a roll of a die is 1/6 = 0.1666667. As above we can use **R** to simulate an experiment of rolling a die a number of times and compare our results with the theoretical probability. Let us now tell R to roll a die 30 times:

```{r}
sample(1:6,30,rep=T)
```
As before one can write a function to roll a die $n$ times:

```{r}
RollDie=function(n) sample(1:6,rep=T)
d1=RollDie(50)
d1
```
Now use the `sum` command to compare the results from this experiment to the theoretical probabilities. For instance, in the above experiment the number of 4's and its relative frequency is:

```{r}
sum(d1==2)
sum(d1==2)/50
```

The number 2 occurs 8 times and its relative frequency is 0.04 which is far away 1/6. Note that you may get different answers. We can plot a relative histogram using the command:

```{r}
hist(d1,breaks=c(0.5,1.5,2.5,3.5,4.5,5.5,6.5), prob=T)
```

## Questions

1. Use R to simulate an experiment of rolling a die 100 times. Print the relative histogram and write your name on it.

2. Use R to roll a die 1000 times. (Do not print histogram)

# Review of Discrete Random variables

For any discrete random variable $X$,

\[\{x,\,\mathbb{P}(X=x)\}\,\,\text{Probability distribution}\]

\[\mathbb{E}(X)=\sum_x x\mathbb{P}(X=x)\,\,\text{Expected value of mean}\]

\[\mathbb{E}(f(X))=\sum_x f(x)\mathbb{P}(X=x)\,\,\text{Expected value of function}\]

\[\text{Var}(X)=\mathbb{E}X^2-(\mathbb{E}X)^2\]

## Combinatorial coefficients

\[{N\choose n}=\frac{N!}{n!(N-n)!},\quad \text{where}\,\,N!=N(N-1)(N-2)\cdots 1\]

## Hypergeometric Distribution

\[\mathbb{P}(\text{Select }n_i\,\,\text{balls of each type }i)=\mathbb{P}(X_1=n_1,\cdots,X_r=n_r;N,n,N_1,\cdots,N_r)=\frac{{N_1\choose n_1}{N_2\choose n_2}\cdots {N_r\choose n_r}}{{N\choose n}}\]
where $N=\sum^r_{i=1}N_i$ and $n=\sum^r_{i=1}n_i$.



